\section{Limitations}
\label{sec:limitations}

We documented seventeen limitations (L1--L17); Table~\ref{tab:limitations} lists the structural
ones with their proposed fixes. Four deserve emphasis.

\textbf{Short episodes (L1)} are the dominant data constraint. At $\sim$21 steps, \texttt{ewat\_v3}
leaves too little window for the look-through warm-up, which is the proximate cause of H2a's
failure, H2b's triviality, and the high false-alarm rate at low thresholds. \texttt{ewat\_v4}
(47--51 steps) mitigates but does not eliminate it (H2a still fails).

\textbf{Evaluation circularity (L9)} is the methodological core of this paper. The
$0.987$ precursor AUROC targets EWAT's own labels; we therefore demote it and adopt the independent
Chaos~Mesh target ($0.920$) as the headline, with the stress tests of Section~\ref{sec:validity} as
the evidence.

\textbf{Type-count instability (L10).} The siamese head overfits in $2$--$7$ epochs on
\texttt{ewat\_v4} and $K$ ranges from $9$ to $15$ across seeds; this is structural (small training
set, diverse episode lengths), not a mining bug. We recommend fixing $K$ or moving to density-based
clustering in future iterations.

\textbf{Scope and external validity (L13, L14, L17).} All experiments use one six-service topology
on one cluster. Zero-shot transfer to RCAEval recovers generic anomaly detection (H1 silhouette
$0.684$) but not type discrimination (H3 $\approx 0.5$); the bottleneck is a non-transferable
scaler. The ontology, while HermiT-consistent and literature-anchored, has not been audited by an
operational SRE, and its causal edges are extracted from synthetic cascades.

\begin{table}[t]
\centering
\caption{Structural limitations and proposed fixes.}
\label{tab:limitations}
\small
\begin{tabularx}{\columnwidth}{@{}lXl@{}}
\toprule
\textbf{ID} & \textbf{Limitation} & \textbf{Fix} \\
\midrule
L1  & short episodes ($\sim$21 steps)      & \texttt{v4}/\texttt{v5} long episodes \\
L9  & H3 circularity                        & independent target \\
L10 & siamese overfit / $K$ unstable        & fixed $K$ / HDBSCAN \\
L13 & $N=6$, one topology                   & Train~Ticket (\texttt{v5}) \\
L14 & no cross-cluster validation           & few-shot domain adapt. \\
L17 & ontology not SRE-audited              & expert review \\
\bottomrule
\end{tabularx}
\end{table}

We also note minor caveats for transparency: a single train/val/test split (no $k$-fold), no a
priori power analysis (post-hoc: $5/10$ clusters are statistically reportable at $n_{\text{pos}}\geq
5$), and stress tests A1--A5 designed \emph{post hoc} rather than pre-registered.
